Awareness and Experience of Developers to\\ Outdated and License-violating Code on Stack Overflow:\\ The Online Survey

### 2017-09-16 cloverflow_technical_report/cloverflow_survey_rn.tex
Latest update on the technical report.

\begin{enumerate}
	\item \textbf{RQ1 (Sources of Stack Overflow Answer Snippets):} \textit{Where are the code snippets in Stack Overflow answers from?}
	\item \textbf{RQ2 (Awareness of Stack Overflow answerers to outdated code):} \textit{Are Stack Overflow answerers aware of outdated code in their answers?}
	\item \textbf{RQ3 (Awareness of Stack Overflow answerers to outdated code):} \textit{Are Stack Overflow answerers aware of outdated code in their answers?}
	\item \textbf{RQ4 (Problems from Stack Overflow code snippets):} \textit{What are the problems Stack Overflow visitors experiencing from reusing code snippets on Stack Overflow?}
	\item \textbf{RQ5 (Software licensing issues from Stack Overflow code snippets):}\textit{How often do Stack Overflow visitors check for software licensing issues from code snippets on Stack Overflow?}
\end{enumerate}

@@

\section{Results and Discussions}
We collected and analysed the results after we closed the surveys on 12
September 2017. We now discuss the results separately for the answerer and the
visitor survey.

@@

The participants are active users and regularly answer questions on Stack
Overflow (see Figure~\ref{fig:survey_answer_frequency}). Eighty-two (82\%) and
sixty (60\%) of answerers from Group 1 and Group 2  answer questions at
least once a week.
More than half of the answerers very frequently (81--100\% of the time) or
frequently (61--80\% of the time) include code snippets in their answers. To
break down into two groups as depicted in
Figure~\ref{fig:survey_answer_frequency_with_code}, Group 1 very frequently
(48.7\%) and frequently (27.4\%) provide code examples when answering. Likewise,
Group 2 does the same (very frequently for 32.1\% and frequently for 36.9\%).

@@

To answer this research question, we ask the participants for original sources
of their code examples. We provided six options (allowing more than one answer)
including \textit{I copied them from my own personal projects}, \textit{I copied
	them from my company's projects}, \textit{I copied them from open source
	projects}, \textit{I wrote the new code from scratch}, \textit{I copied the code
	from the question and modify it for the answer}, and \textit{Others}. The
answers are shown in Figure~\ref{fig:survey_snippet_source}. Participants in
Group 1 mainly write new code from scratch (116) or copy from the code snippets
in question and modify it for the answer (112), followed by from their personal
projects (105), open source projects (77), other sources (59), and company
projects (48). For Group 2, the main source is also writing code from scratch
(83), followed by copying from personal projects and modifying from the question
(77), copying from open source projects (56), copying from other sources (40),
and copying from company projects (31).

@@

Half of the top answerers on Stack Overflow
are aware of outdated code in their answers. We found that 71.6\% of Stack Overflow answerers in Group 1 have been notified of
outdated code in at least one of their answers (see Figure~\ref{fig:survey_outdated}). The ratio drops to 57.1\% in
Group 2.

@@

We then asked the participants who have been notified of their outdated code
(81 and 48 participants from Group 1 and 2 respectively) a
follow-up questions \textit{``how frequently did you fix your outdated code on Stack
Overflow?''} The answers, depicted in Figure~\ref{fig:survey_outdated_fix}, show that
more than half of them frequently fix the outdated code snippets. However, there
are 19.8\% and 18.8\% of the answerers in Groups 1 and 2 who rarely or never fix their code.

@@

As shown in Figure~\ref{fig:survey_license_known}, more than half of the
answerers in both groups (62.1\% and 61.9\% respectively) are aware that Stack
Overflow apply Creative Commons Attribution-ShareAlike 3.0 Unported (CC BY-SA
3.0) to content in the posts, including code snippets, while the rest of 37.9\%
and 38.1\% are not.

@@

Almost every answerer in both groups (98\% and 99\% respectively) does not
include license statement in their code snippets (numbers are shown in
Figure~\ref{fig:survey_license}). Some of the participants explain the reason in
the open comment question which we summarised into three groups as follows.
First, they choose to post only their own code or code that is adapted from the
question. The code is automatically subjected to Stack Overflow's CC BY-SA 3.0.
Second, they copy the code from company or open source projects that they know are
permitted to be publicly distributed. Hence, no license statement is required.
Third, some answerers believe that code snippets in their answers are too small
to claim any intellectual property on them and fall under fair
use~\citep{fairuse}.

@@

While nobody explicitly includes a software license in their snippets,
many users include a statement on their profile page that states all their
answers are under a certain license. For example, \textit{All code posted by me on
	Stack Overflow should be considered public domain without copyright. For
	countries where public domain is not applicable, I hereby grant everyone the
	right to modify, use and redistribute any code posted by me on Stack Overflow
	for any purpose. It is provided "as-is" without warranty of any kind.}  Many
users either declare their snippets to be public domain, or they grant
additional licenses, e.g. Apache 2.0 or MIT/Expat.

@@

We asked the answerers a follow-up question of how frequently they checked for a
conflict between software license of the code snippets they copied to their
answers and Stack Overflow's CC BY-SA 3.0. As shown in
Figure~\ref{fig:survey_license_check}, 69\% of answerers from both groups did
not perform the checking. There are only approximately 10\% of the answerers who
frequently checks for licensing conflicts when they copy code snippets to Stack
Overflow.

\vspace{0.5cm} \noindent\fbox{%
	\parbox[c][1.5cm]{0.98\textwidth}{%
		\textit{For RQ3, approximately 60\% of answerers are aware of CC BY-SA 3.0
			license enforced by Stack Overflow. However, 98--99\% of the answerers do not
			include software license in their Stack Overflow snippets nor check for
			licensing conflicts when they copy code snippets to Stack Overflow answers.}
	}%
} \vspace{0.5cm}

In addition, we invited the answerers to give comments regarding their concerns
when answering Stack Overflow (SO) with code snippets. Some interesting comments
are selected and discussed below. The full set of answers can be found in the
Appendix~\ref{appendixC}.

@@

	\item Comment 4: The answerer discusses a problem from outdated Stack Overflow code snippets and
	his/her solution.

@@

To answer RQ4 and RQ5, we ask Stack Overflow visitors about their experiences of
outdated code and their awareness to software license of Stack Overflow code
snippets. We received 89 answers from 5 groups of Stack Overflow visitors. The
combined results and discussions are presented below.

@@

\begin{figure}
		\centering
		\includegraphics[width=0.5\linewidth]{survey_visitor_exp}
		\caption{Experience of the Stack Overflow visitors}
		\label{fig:survey_visitor_exp}
\end{figure}

@@

\subsubsection*{RQ4: Why developers choose to find programming solutions on Stack Overflow?}

Similar to a study by~\cite{Acar2016}, we asked the participants to rank five
options, without a tie, they will choose to find programming solutions. The
given five options include books, official documentations, Stack Overflow,
online repositories (e.g. GitHub), and others. The results are displayed in
Figure~\ref{fig:survey_visitor_rankings-crop}. We found that 47 out of 89
participants rank Stack Overflow as the 1st option to search for programming
solutions, followed by official documentation (33), online repositories (6),
other resources (5), and books (1).

@@

We then shift the focus to only Stack Overflow. We ask the participants how
frequently they reuse code snippets from answers on Stack Overflow. According to
the answers (see Figure~\ref{fig:survey_visitor_frequency_so_copy-crop}), we
found that 57 (64\%) participants actively reusing code snippets from Stack
Overflow. Eight participants (9\%) copy Stack Overflow code every day, 22 (25\%)
do copying 3-6 times a week, 27 (30\%) do copying once or twice a week. There
are 2 participants (2\%) who never copy the code from Stack Overflow.

@@

To understand why the participants choose to copy code snippets from Stack
Overflow, we asked them to rate four reasons in Likert's scale (Strongly agree,
Agree, Undecided, Disagree, Strongly disagree). The four reasons include
\textit{They are easy to find by searching the web}, \textit{They solve problems
	similar to my problems with minimal changes}, \textit{The context of questions
	and answers helped me understand the code snippets better}, \textit{The voting
	mechanism and accepted answers helped to filter good code from bad code}. The
answers are depicted in Figure~\ref{fig:survey_visitor_why_copy_so}. More than 80\% of the
participants agree with all the four reasons.

@@

\vspace{0.5cm} \noindent\fbox{%
	\parbox[c][1.5cm]{0.98\textwidth}{%
		\textit{For RQ4, Stack Overflow is ranked higher than official documentations,
	online repositories, and book as the resource to look for programming
	solutions. Developers rely on Stack Overflow answers because they are easy to
	search for on the web. Moreover, 64\% of the participants reuse code snippets
	from Stack Overflow at least once a week.}
	}%
} \vspace{0.5cm}

>>>>

\begin{enumerate}
	\item \textbf{RQ1 (Sources of Stack Overflow Answer Snippets):} \textit{Where are the code snippets in Stack Overflow answers from?}
	\item \textbf{RQ2 (Awareness of Stack Overflow answerers to outdated code):} \textit{Are Stack Overflow answerers aware of outdated code in their answers?}
	\item \textbf{RQ3 (Awareness of Stack Overflow answerers to outdated code):} \textit{Are Stack Overflow answerers aware of outdated code in their answers?}
	\item \textbf{RQ4 (Search for Coding Solutions):} \textit{Where do the developers search for programming solutions?}
	\item \textbf{RQ4 (Problems from Stack Overflow code snippets):} \textit{What are the problems Stack Overflow visitors experiencing from reusing code snippets on Stack Overflow?}
	\item \textbf{RQ5 (Software licensing issues from Stack Overflow code snippets):}\textit{How often do Stack Overflow visitors check for software licensing issues from code snippets on Stack Overflow?}
\end{enumerate}

@@

\section{Results and Discussions}
We collected and analysed the results after we closed the surveys on 12
September 2017. We now discuss the results from the answerer and
visitor survey separately.

@@

The participants are active users and regularly answer questions on Stack
Overflow (see Figure~\ref{fig:survey_answer_frequency}). Eighty-two (82\%) and
sixty (60\%) of answerers from Group 1 and Group 2  answer questions at least
once a week. More than half of the answerers very frequently (81--100\% of the
time) or frequently (61--80\% of the time) include code snippets in their
answers. To break down into two groups as depicted in
Figure~\ref{fig:survey_answer_frequency_with_code}, Group 1 very frequently
(48.7\%) and frequently (27.4\%) provide code examples when answering. Likewise,
Group 2 follow the same trend (very frequently for 32.1\% and frequently for
36.9\%). Interestingly, there is one participant who never include code snippet
in his/her answer thus the results after this question are from 116 participants.

@@

To answer this research question, we asked the participants for the original
source of their code examples. We provided six options (allowing more than one
answer) including \textit{I copied them from my own personal projects},
\textit{I copied them from my company's projects}, \textit{I copied them from
	open source projects}, \textit{I wrote the new code from scratch}, \textit{I
	copied the code from the question and modify it for the answer}, and
\textit{Others}. The answers are shown in
Figure~\ref{fig:survey_snippet_source}. Participants in Group 1 mainly write new
code from scratch (116) or copy from the code snippets in question and modify it
for the answer (112), followed by from their personal projects (105), open
source projects (77), other sources (59), and company projects (48). For Group
2, the main source is also writing code from scratch (83), followed by copying
from personal projects and modifying from the question (77), copying from open
source projects (56), copying from other sources (40), and copying from company
projects (31). There are 133 answerers out of total 607 from two groups who have cloned
code snippets from open source projects into their answers at least once. We are interested
in this type of clones and will investigate further in the later RQs.

@@

Half of the top answerers on Stack Overflow are aware of outdated code in their
answers. As visualised in Figure~\ref{fig:survey_outdated}, 83 participants (71.6\%)
of Stack Overflow answerers in Group 1 have been notified of outdated code in at
least one of their answers. The ratio drops to 48 participants (57.1\%) in Group 2.

@@

We then asked the participants who have been notified of their outdated code (83
and 48 participants from Group 1 and 2 respectively) a follow-up questions
\textit{``how frequently did you fix your outdated code on Stack Overflow?''}
The answers, depicted in Figure~\ref{fig:survey_outdated_fix}, show that more
than half of them frequently fix the outdated code snippets. However, there are
17 (19.8\%) and 9 (18.8\%) participants in Groups 1 and 2 who rarely, very
rarely or never fix their code.

Regarding this issue of outdated code, one participant expresses their concern
in the open comment question: \textit{``The main problem for me/us is outdated
	code, esp. as old answers have high Google rank so that is what people see
	first, then try and fail. Thats why we're moving more and more of those examples
	to knowledge base and docs and rather link to those.''} On the other hand,
another participant does not  worry about his/her outdated code and how he
or she handles them: \textit{``On
	the matter of deprecation, I almost entirely use .NET which has got different
	versions of the framework. Therefore, code deprecation is not often a problem
	since what is deprecated on one version of the framework may be the only way of
	solving a given problem on an older version of the framework. I may also have to
	add that questions I tend to answer are about how to solve general coding
	problems so they are not usually subject to deprecation.''}

@@

As shown in Figure~\ref{fig:survey_license_known}, more than half of the
answerers in both groups, 72 (62.1\%) and 53 (62.3\%) respectively, are aware that Stack
Overflow apply Creative Commons Attribution-ShareAlike 3.0 Unported (CC BY-SA
3.0) to content in the posts, including code snippets, while the rest of 44 (37.9\%)
and 32 (37.6\%) are not.

@@

Almost every answerer in both groups, 114 out of 116 (98\%) and 84 out of 85
(99\%) respectively, does not include license statement in their code snippets
(numbers are shown in Figure~\ref{fig:survey_license}). Some of the participants
explain the reason in the open comment question which we summarised into three
groups as follows. First, they choose to post only their own code or code that
is adapted from the question. The code is automatically subjected to Stack
Overflow's CC BY-SA 3.0 without any explicit licensing statement. Second, they
copy the code from company or open source projects that they know are permitted
to be publicly distributed. Hence, no license statement is required. Third, some
answerers believe that code snippets in their answers are too small to claim any
intellectual property on them and fall under fair use~\citep{fairuse}.

@@

While almost nobody explicitly includes a software license in their snippets,
many participants include a statement on their profile page that all their
answers are under a certain license. For example, \textit{All code posted by me on
	Stack Overflow should be considered public domain without copyright. For
	countries where public domain is not applicable, I hereby grant everyone the
	right to modify, use and redistribute any code posted by me on Stack Overflow
	for any purpose. It is provided "as-is" without warranty of any kind.}  Many
participants either declare their snippets to be public domain, or they grant
additional licenses, e.g. Apache 2.0 or MIT/Expat.

@@

We asked the answerers a follow-up question of how frequently they checked for a
conflict between software license of the code snippets they copied to their
answers and Stack Overflow's CC BY-SA 3.0. As shown in
Figure~\ref{fig:survey_license_check}, 80 (69\%) and 58 (69\%) answerers from
group 1 and group 2 did not perform the checking. There are only approximately
10\% of the answerers who frequently checks for licensing conflicts when they
copy code snippets to Stack Overflow.

\vspace{0.5cm} \noindent\fbox{%
	\parbox[c][1.5cm]{0.98\textwidth}{%
		\textit{For RQ3, approximately 62\% of our participants are aware of CC BY-SA
			3.0 license enforced by Stack Overflow. However, 98--99\% of the answerers
			never include software license in their Stack Overflow snippets. Sixty-nine
			percent never check for licensing conflicts when they copy code snippets to
			Stack Overflow answers.} }%
} \vspace{0.5cm}

To acquire additional insights, we invited every answerers for
further comments regarding their concerns of answering Stack Overflow (SO)
with code snippets. Some interesting comments are selected and discussed below.
The full set of answers can be found in the Appendix~\ref{appendixC}.

@@

	\item Comment 4: As shown in RQ2, the answerer discusses a problem from outdated Stack Overflow code snippets and
	his/her solution.

@@

To answer RQ4 and RQ5, we used another online survey, the visitor survey, to ask
Stack Overflow visitors about their experiences of obsolete code and their
awareness to software license of Stack Overflow code snippets. We received 89
answers from 5 groups of Stack Overflow visitors. We combined the results and
present them as a single group below.

@@

\begin{figure}
		\centering
		\includegraphics[width=0.4\linewidth]{survey_visitor_exp}
		\caption{Experience of the Stack Overflow visitors}
		\label{fig:survey_visitor_exp}
\end{figure}

@@

\subsubsection*{RQ4: Where do the developers search for programming solutions?}

We asked the participants to rank five options, without a tie, they will choose
to find programming solutions. The given five options include books, official
documentations, Stack Overflow, online repositories (e.g. GitHub), and others.
The results are displayed in Figure~\ref{fig:survey_visitor_rankings-crop}. We
found that 47 out of 89 participants rank Stack Overflow as the 1st option to
search for programming solutions, followed by official documentation (33),
online repositories (6), other resources (5), and books (1).

@@

Since Stack Overflow is among the first research to solve programming tasks, we
ask the participants how frequently they reuse code snippets from answers on
Stack Overflow. According to the results (see
Figure~\ref{fig:survey_visitor_frequency_so_copy-crop}), we found that 57 (64\%)
participants actively reusing code snippets from Stack Overflow. Eight
participants (9\%) copy Stack Overflow code every day, 22 (25\%) do copying 3-6
times a week, 27 (30\%) do copying once or twice a week. There are 2
participants (2\%) who never copy the code from Stack Overflow.

@@

To understand why the participants choose to copy code snippets from Stack
Overflow, we asked them to rate four reasons in Likert's scale (Strongly agree,
Agree, Undecided, Disagree, Strongly disagree). The four reasons include
\textit{They are easy to find by searching the web}, \textit{They solve problems
	similar to my problems with minimal changes}, \textit{The context of questions
	and answers helped me understand the code snippets better}, \textit{The voting
	mechanism and accepted answers helped to filter good code from bad code}. The
answers are depicted in Figure~\ref{fig:survey_visitor_why_copy_so}. More than
80\% of the participants agree with all the four reasons. We observed only two
``disagree'' and zero ``Strongly disagree'' answer for ``Helpful context'', the
lowest disagreement among the four reasons. This means most of them agree that
the context of questions and answers on Stack Overflow help them understand the
code snippet better.

@@

\vspace{0.5cm} \noindent\fbox{%
	\parbox[c][2cm]{0.98\textwidth}{%
		\textit{For RQ4, Stack Overflow is ranked higher than official documentations,
	online repositories, and book as the resource to look for programming
	solutions. Developers rely on Stack Overflow answers because they are easy to
	search for on the web. Moreover, 64\% of the participants reuse code snippets
	from Stack Overflow at least once a week. The copy code from Stack Overflow
    because they can be found easily from search engine, solve similar problems to
    their problems, provide helpful context, and offer voting and accepted answers.}
	}%
} \vspace{0.5cm}
